Towards a Practical Face Recognition System: \\ Robust Alignment and Illumination by Sparse Representation

### 2010-12-31 rebuttal.tex

\item Rigorous justification of the choice of illuminations is a subject of
continuing research.  Our current choice is motivated empirically by the
experiments described in the paper.

\item To our knowledge, we are the first to use projectors for illumination in
face recognition applications.  Section 1A of the previous submission already
cites several other applications that use projectors as illumination sources.
The requested references propose a technique for improving the SNR of projector-based
illumination systems for certain classes of objects.  We have added
discussion of this to the paper.

\item  It is true that one or both stages of the algorithm could be substituted
with other techniques.  That said, the registration and recognition stages are
coherent in the sense that they are motivated by same modeling assumptions
(illumination model, sparse occlusion, etc.), they use the same robust penalty
on image error (the L1-norm), and they use the same numerical optimization
techniques (ALM).  While conceptual simplicity is a secondary concern in
algorithm design, starting with a minimalist foundation benefits both
practitioners as well as researchers.  Other researchers have already published
more complex extensions to our work.

\item  There is no extensive computation involved in
    Step 13 of Algorithm 1. In fact, for each candidate
    subject $i\in {k_1, \ldots, k_s}$ for recognition, Step
    13 simply uses the transformation $\tau_i$ obtained
    from the iterative alignment stage (Step 2-9) to
    re-sample each original training
    image of that subject, so that this new set of training
    samples is aligned with the test sample. Nevertheless,
    it worth mentioning that, since we perform alignment
    separately for each class in the gallery, the
    computation for the alignment stage (Step 2-9) is
    linear in the size of the gallery. However, thanks to
    its separability, the alignment stage can be easily
    parallelized.

@@

\item (iii) ``How often the solution gets drifted away into a wrong basin of
attraction. (If there is any way to know this.)''

In Section II E, we have conducted careful experiments to measure the region of attraction
for the proposed alignment method. We have studied the region of attraction of the proposed
method for both 2D and 3D pose variation, as well as how it can be improved with a proper
multi-scale implementation.

\item (iv) ``What is the criteria or threshold for convergence? If the threshold is
stringent, the algorithms are likely to get stuck in infinite loops, and if the
threshold is liberal, premature convergence would occur. How are such
thresholds picked and how universal are they? Does every new dataset need its
own threshold?''

For the $\ell_1$ minimizations we terminate the ALM
algorithm when the relative changes of primal variables in
two consecutive iterations become small, as suggested by
[27]. That is,
\begin{displaymath}
\frac{\|\x_{k+1}-\x_k\|_2}{\|\x_k\|_2} < \epsilon \quad
\textup{and} \quad \frac{\|\e_{k+1}-\e_k\|_2}{\|\e_k\|_2} <
\epsilon,
\end{displaymath}
where $\epsilon>0$ is the tolerance. With the suggested
criteria, the l1-minimizations are guaranteed to converge
eventually, although convergence may be very slow if they
are chosen poorly. For the alignment
iteration, we terminate when the change of objective value
(\ie $\|\e\|_1$) in two consecutive iterations becomes
small. We use the same set of convergence thresholds for
all the data sets we ran against.

@@

\end{enumerate}

\section{Response to Reviewer \#3}
Reviewer 3 makes the following requests:

\begin{enumerate}
\item ``I would have liked to see the experiment for the case S=1, without
impostors in the dataset, to judge the effect of the ``Sparse
Concentration Index''.

We have added experiments to the paper demonstrating
the utility of the sparse representation(SR)-based
recognition stage for classification.  In general, we see a
few percentage points of improvement in recognition rate
when using the representation residual obtained from the
SR-based recognition stage (with $S=10$) over the
recognition based purely on alignment error (with $S$=1).
While this effect is desirable, as discussed in the paper,
it is worthy emphasizing that the SR-based recognition also
enables us to compute SCI, which has demonstrated its power
for impostor rejection.

### 2011-01-04 faces_system.tex
Worked on introduction

\PARstart{T}{here} has been a historical tendency for face recognition algorithms to work well under laboratory conditions  (well aligned images, similar lighting conditions, and little change in appearance) but degrade drastically when tested in less-controlled real-world environments. Although achieving high recognition performance under uncontrolled conditions and with limited gallery images has been the traditional goal of face recognition research, existing techniques are not yet able to meet the performance requirements of high-stakes security-related applications. While there have been many successes in less-demanding applications such as online image search and family photo album organization (e.g.\ Google Picassa, Microsoft Photo Gallery, and Apple iPhoto), there have been a number of high-profile failed trials of face recognition technology in extremely challenging mass surveillance/watch-list applications.

Somewhat surprisingly, face recognition has not seen widespread applications or success in more tractable security settings, such as access control for secure facilities (e.g., prisons, office buildings), computer systems, automobiles, or automatic teller machines, where controlled gallery images can be obtained in advance.
%\footnote{Face recognition scenarios where both gallery and test data are uncontrolled continue to be an active research area \cite{LFW}, and has become more popular recently due to the aforementioned online applications.}  
These applications are very interesting due to their potential sociological and technological impact.  They are also qualitatively different from mass surveillance or online image applications because they demand extremely reliable systems whose performance must compete with other biometric measures such as fingerprint and iris. Because the gallery subjects are allies, rather than opponents, of the recognition system, this creates the possibility of carefully controlling the acquisition of the training data (as is common practice for other biometrics). However, unlike other biometrics, face recognition has the potential of working with test data that is much less controlled, allowing the access control system to be made more discreet and much less intrusive. To some extent, the goal of this paper is to show how a reliable face recognition system can be built for this admittedly restrictive, but still important, scenario.

Classical holistic subspace-based face recognition methods \cite{Turk1991-CVPR,Belhumeur1997-PAMI} are well known for their speed and simplicity, as well as for their natural extension to linear illumination
models.  However, their performance has been shown to be extremely brittle not
only to alignment variation, but to even minor occlusions caused by, say, a
wisp of hair, a blinked eye, or mouth that is slightly open. One promising
recent direction, set forth in \cite{Wright2009-PAMI}, casts the recognition
problem as one of finding a sparse representation of the test image in terms of
the training set as a whole, up to some sparse error due to occlusion. A
\emph{sparse representation-based classification} (SRC) method is then proposed
for recognition. The main idea is that the sparse nonzero coefficients should
concentrate on the training samples with the same class label as the test
sample. SRC has demonstrated striking recognition performance despite severe
occlusion or corruption by solving a simple convex program.  
%For this reason, the final recognition stage of our proposed system will consist of SRC,
%and the alignment stage that precedes it will be based on similar techniques.

>>>>

\PARstart{F}{ace} recognition applications to date have fallen into roughly two
categories.  Face recognition has recently seen a lot of success in a family of
less-demanding applications such as online image search and family photo album
organization (e.g.\ Google Picassa, Microsoft Photo Gallery, and Apple iPhoto).
At the other end of the tractability spectrum there are the terrorist watchlist
and mass surveillance applications that have for the most part dominated the
field of face recognition research.  However, there are many face recognition
that fall roughly between these extremes, where very high recognition
performance is desired, but the users in the gallery are still allies of the
system rather than adversaries.  These applications include access control for
secure facilities (e.g., prisons, office buildings), computer systems,
automobiles, or automatic teller machines, where controlled gallery images can
be obtained in advance.  These applications are very interesting due to their
potential sociological impact.  Since the gallery subjects are allies, rather
than opponents, of the recognition system, this creates the possibility of
carefully controlling the acquisition of the training data. While the same can
be said for other biometrics such as fingerprints and iris recognition, face
recognition has the potential of working with test data that is much less
controlled, allowing the access control system to be made less intrusive to the
users of the system. To some extent, the goal of this paper is to show how a
reliable face recognition system can be built for this admittedly restricted,
but still important, scenario.

Very few recognition systems specifically target applications where many well-
controlled training images are available.  Of these, the classical holistic
subspace-based face recognition methods \cite{Turk1991-CVPR,Belhumeur1997-PAMI}
are well known for their speed and simplicity, as well as for their natural
extension to linear illumination models.  However, their performance has been
shown to be extremely brittle not only to alignment variation, but to even
minor occlusions caused by, say, a wisp of hair, a blinked eye, or mouth that
is slightly open. One promising recent direction, set forth in
\cite{Wright2009-PAMI}, casts the recognition problem as one of finding a
sparse representation of the test image in terms of the training set as a
whole, up to some sparse error due to occlusion. A \emph{sparse
representation-based classification} (SRC) method is then proposed for
recognition. The main idea is that the sparse nonzero coefficients should
concentrate on the training samples with the same class label as the test
sample. SRC has demonstrated striking recognition performance despite severe
occlusion or corruption by solving a simple convex program.  
%For this reason, the final recognition stage of our proposed system will
%consist of SRC, and the alignment stage that precedes it will be based on
%similar techniques.
